---
title: "Padula_2025_Wing_Diff_Analysis"
author: "Sara Padula"
date: "2025-03-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

```{r}
library(dplyr)
library(ggplot2)
library(cowplot)
library(lme4)
library(lmerTest)
library(DHARMa)
library(glmmTMB)
library(car)
library(emmeans)
library(ggpubr)
library(ggdist)
library(patchwork)
```

## Load data


```{r}
#Loading data
breed <- read.csv("paired_SMI.csv")

## Scaling elevation variable for each species
breedMOCH <- breed %>%
  filter(Species == "MOCH") %>%
  mutate(scale_ele = scale(Elevation))

breedBCCH <- breed %>%
  filter(Species == "BCCH") %>%
  filter(Elevation < 2000) %>%
  mutate(scale_ele = scale(Elevation))

breed <- rbind(breedMOCH, breedBCCH)

#What is the sample size and mean wing difference in...

#MOCH?
nrow(breedMOCH)

mean(breedMOCH$Wing_Difference)


#BCCH?
nrow(breedBCCH)

mean(breedBCCH$Wing_Difference)

```

## Breeding Success Analysis -- Numerical wing difference variable
```{r}
#first egg, removing blank values
breed.fe <- breed %>% filter(!is.na(First.Egg))
sum(breed.fe$Species == "MOCH") 
sum(breed.fe$Species == "BCCH")

#clutch size, removing blank values
breed.cs <- breed %>% filter(!is.na(Egg_Number))
sum(breed.cs$Species == "MOCH") 
sum(breed.cs$Species == "BCCH")

#brood size, removing blank values
breed.bs <- breed %>% filter(!is.na(Nestling_Number))
sum(breed.bs$Species == "MOCH") 
sum(breed.bs$Species == "BCCH")

#nestling size, removing blank values
breed.mm <- breed %>% filter(!is.na(Avg_Nestling_Weight))
sum(breed.mm$Species == "MOCH") 
sum(breed.mm$Species == "BCCH")

#Female SMI, removing blank values
breed.fs <- breed %>% filter(!is.na(Female_SMI))
sum(breed.fs$Species == "MOCH") 
sum(breed.fs$Species == "BCCH")

#Male SMI, removing blank values
breed.ms <- breed %>% filter(!is.na(Male_SMI))
sum(breed.ms$Species == "MOCH") 
sum(breed.ms$Species == "BCCH")

#Provisioning
breed.p <- breed %>% filter(!is.na(Provisioning))
sum(breed.p$Species == "MOCH") 
sum(breed.p$Species == "BCCH")


```

## Modeling date of first egg

```{r}
# Centering within year
breed.fe$First.Egg_YearCentered <- ave(breed.fe$First.Egg, breed.fe$Year, FUN = function(x) scale(x, center = TRUE, scale = FALSE))

# Centering within elevation (after year)
breed.fe$First.Egg_Scaled <- ave(breed.fe$First.Egg_YearCentered, breed.fe$scale_ele, FUN = function(x) scale(x, center = TRUE, scale = FALSE))

breed.fe$scale_ele <- as.numeric(breed.fe$scale_ele)

## Model first egg ######
fe1 <- lmer(First.Egg_Scaled ~ Wing_Difference*Species + scale_ele + (1|Year), data = breed.fe)
fe2 <- lmer(First.Egg_Scaled ~ Species + scale_ele + (1|Year), data = breed.fe)

anova(fe2, fe1) #wing difference doesn't significantly improve model fit

#Check residuals
fe1r = simulateResiduals(fe1)
plot(fe1r)
#quantile deviation

#wing diff doesnt improve fit and the model deviates, so we can confidently say that wing difference within pairs doesn't impact date of first egg.

summary(fe1) # species significant

mochFEp <- 0.4817

bcchFEp <- 0.2326


```

## Modeling clutch size

```{r}
cs1 <- glmmTMB(Egg_Number ~ Wing_Difference*Species, data =breed.cs,family = "genpois")
cs2 <- glmmTMB(Egg_Number ~ Species, data=breed.cs, family="genpois") 

anova(cs1,cs2) # wing difference significantly improves model fit

#Check residuals
cs1r = simulateResiduals(cs1)
plot(cs1r)
#outliers significant on q-q plot

summary(cs1) #species significant

mochCSp <- 0.101

bcchCSp <- 0.795


```

## Modeling brood size

```{r}
bs1 <- glmmTMB(Nestling_Number ~ Wing_Difference*Species + (1|Year), data = breed.bs, family="genpois")
bs2 <- glmmTMB(Nestling_Number ~ Species + (1|Year), data = breed.bs, family = "genpois") #got error when including (1|Year) and variance of the random effect was very low

anova(bs1,bs2) #wing difference significantly improves model fit

#check residuals
bs1r = simulateResiduals(bs1)
plot(bs1r)
#look good

#Results
summary(bs1) #species and wing_difference*species significant

mochBSp <- 0.00118

bcchBSp <- 0.353


```

## Modeling average nestling mass

```{r}
##Model nestling mass #####
mm1 <- lmer(Avg_Nestling_Weight ~ Wing_Difference*Species + Nestling_Number + (1|Year), data = breed.mm)
mm2 <- lmer(Avg_Nestling_Weight ~ Species + Nestling_Number + (1|Year), data = breed.mm)

anova(mm1, mm2) # wing difference significantly improves model fit

#check residuals
mm1r = simulateResiduals(mm1) 
plot(mm1r)
#some deviation

#results
summary(mm1) #wing_difference*species significant

mochMMp <- 0.0204

bcchMMp <- 0.7428
```

## Modeling Female SMI

```{r}
fs1 <- lmer(Female_SMI ~ Wing_Difference*Species + (1|Year), data =  breed.fs)
fs2 <- lmer(Female_SMI ~ Species + (1|Year), data = breed.fs)

anova(fs1,fs2) # wing difference doesn't improve model fit

#check residuals
fs1r = simulateResiduals(fs1)
plot(fs1r)
#look okay

#results
summary(fs1) #species*elevation significant, elevation almost significant

mochFSp <- 0.990

bcchFSp <- 0.566

```

## Modeling male SMI

```{r}
ms1 <- lmer(Male_SMI ~ Wing_Difference*Species + (1|Year), data =  breed.ms)
ms2 <- lmer(Male_SMI ~ Species + (1|Year), data = breed.ms)

anova(ms1,ms2) #wing difference significantly improves model fit

#check residuals
ms1r = simulateResiduals(ms1)
plot(ms1r)
#look good

#results
summary(ms1) #wing_difference significant

mochMSp <- 0.14301

bcchMSp <- 0.00593
```

## Modeling provisioning rate

```{r}
p1 <- lmer(Provisioning ~ Wing_Difference*Species + (1|Year), data =  breed.p)
p2 <- lmer(Provisioning ~ Species + (1|Year), data = breed.p)

anova(p1,p2) #wing difference doesn't improve model fit

#check residuals
p1r = simulateResiduals(p1)
plot(p1r)
#quantile deviations

#results
summary(p1) #nothing significant

mochPp <- 0.787

bcchPp <- 0.892
```

Adjust the p-values to account for multiple comparisons

```{r}
p_values <- c(mochFEp, bcchFEp, mochCSp, bcchCSp, mochBSp, bcchBSp, mochMMp, bcchMMp, mochFSp, bcchFSp, mochMSp, bcchMSp, mochPp, bcchPp)

p_adjusted <- p.adjust(p_values, method = "BH")

p_adjusted

```


Visualizing wing difference (numeric) predicting reproductive success variables.

```{r}

# Use your fitted model with Wing_Difference
breed.fe$Species <- as.factor(breed.fe$Species)

fe1 <- lmer(First.Egg_Scaled ~ Wing_Difference * Species + scale_ele + (1|Year), data = breed.fe)

# Generate predicted values over range of Wing_Difference for each species
preds_a <- ggpredict(fe1, terms = c("Wing_Difference", "Species"))

# Plot
a <- ggplot(preds_a, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 0.8) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
  labs(x = "", y = "Pred. Scaled Date of First Egg", color = "Species", fill = "Species") +
    scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2"))+
  theme_minimal()+
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 6))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

# Use your fitted model with Wing_Difference
breed.cs$Species <- as.factor(breed.cs$Species)

cs1 <- glmmTMB(Egg_Number ~ Wing_Difference*Species, data =breed.cs,family = "genpois")

# Generate predicted values over range of Wing_Difference for each species
preds_b <- ggpredict(cs1, terms = c("Wing_Difference", "Species"))


b <- ggplot(preds_b, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 0.8) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
  labs(x = "", y = "Pred. Clutch Size", color = "Species", fill = "Species") +
    scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2"))+
  theme_minimal()+
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 6))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

# Use your fitted model with Wing_Difference
breed.bs$Species <- as.factor(breed.bs$Species)

bs1 <- glmmTMB(Nestling_Number ~ Wing_Difference*Species, data =breed.bs,family = "genpois")

# Generate predicted values over range of Wing_Difference for each species
preds_c <- ggpredict(bs1, terms = c("Wing_Difference", "Species"))


c <- ggplot(preds_c, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 0.8) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
  labs(x = "", y = "Pred. Brood Size", color = "Species", fill = "Species") +
    scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2"))+
  theme_minimal()+
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 6))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

breed.mm$Species <- as.factor(breed.mm$Species)

mm1 <- lmer(Avg_Nestling_Weight ~ Wing_Difference*Species + Nestling_Number + (1|Year), data = breed.mm)

# Generate predicted values over range of Wing_Difference for each species
preds_d <- ggpredict(bs1, terms = c("Wing_Difference", "Species"))

d <- ggplot(preds_d, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 0.8) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
  labs(x = "", y = "Pred. Avg. Nestling Mass (g)", color = "Species", fill = "Species") +
    scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2"))+
  theme_minimal()+
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 5))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

breed.fs$Species <- as.factor(breed.fs$Species)

fs1 <- lmer(Female_SMI ~ Wing_Difference*Species + (1|Year), data = breed.fs)

# Generate predicted values over range of Wing_Difference for each species
preds_e <- ggpredict(fs1, terms = c("Wing_Difference", "Species"))

e <- ggplot(preds_e, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 0.8) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
  labs(x = "", y = "Pred. Female Body Condition", color = "Species", fill = "Species") +
    scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2"))+
  theme_minimal()+
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 5))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

breed.ms$Species <- as.factor(breed.ms$Species)

ms1 <- lmer(Male_SMI ~ Wing_Difference*Species + (1|Year), data = breed.ms)

# Generate predicted values over range of Wing_Difference for each species
preds_f <- ggpredict(ms1, terms = c("Wing_Difference", "Species"))


f <- ggplot(preds_f, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 0.8) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
  labs(x = "Wing Difference (mm)", y = "Pred. Male Body Condition", color = "Species", fill = "Species") +
    scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2"))+
  theme_minimal()+
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 5))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

breed.p$Species <- as.factor(breed.p$Species)

p1 <- lmer(Provisioning ~ Wing_Difference*Species + (1|Year), data = breed.p)

# Generate predicted values over range of Wing_Difference for each species
preds_g <- ggpredict(p1, terms = c("Wing_Difference", "Species"))


g <- ggplot(preds_g, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 0.8) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
  labs(x = "Wing Difference (mm)", y = "Pred. Provisioning Rate", color = "Species", fill = "Species") +
    scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2"))+
  theme_minimal()+
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 5))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

allLine <- (a + b + c)/(d + e + f + g)

ggsave("wingdiffRepVarPlot.png", plot = allLine, width = 6, height = 4, dpi = 300)

```


## Does female wing length predict male wing length within pairs?

```{r}
#subsetting data
mochBreed <- subset(breed, Species == "MOCH")
bcchBreed <- subset(breed, Species == "BCCH")

#create new variable for female wing chord

mochBreed$Female.Wing.Chord2 <- mochBreed$Female.Wing.Chord^2

bcchBreed$Female.Wing.Chord2 <- bcchBreed$Female.Wing.Chord^2

#MOCH

#which quadratic model is the best?
head(mochBreed)
quadM1 <- lmer(Male.Wing.Chord ~ Female.Wing.Chord + Female.Wing.Chord2 + (1|Year), data = mochBreed)
linM1 <- lmer(Male.Wing.Chord ~ Female.Wing.Chord + (1|Year), data = mochBreed)

anova(quadM1, linM1) #quadratic term doesnt improve model fit

summary(linM1)

#female wing length doesnt predict male wing length in mountain chickadees

#BCCH

#Does female wing predict male wing in BCCH
#female bander and male bander do not explain any variance so they are excluded. Year does explain variance so it is included.
head(bcchBreed)
quadB1 <- lmer(Male.Wing.Chord ~ Female.Wing.Chord + Female.Wing.Chord2 + (1|Year), data = bcchBreed)
linB1 <- lmer(Male.Wing.Chord ~ Female.Wing.Chord + (1|Year), data = bcchBreed)

summary(linB1)

anova(quadB1, linB1)

#The quadratic model has a slightly lower AIC (better fit), but the likelihood ratio test is not significant (p = 0.11), meaning the addition of the quadratic term does not significantly improve model fit over the linear model.

summary(linB1)

#female wing length doesnt predict male wing length in black-capped chickadees

```

# Plot the linear models
```{r}

ggplot(breed, aes(x = Female.Wing.Chord, y = Male.Wing.Chord, color = Species)) +
  geom_point(alpha = 0.7) +  
  geom_smooth(method = "lm", se = FALSE) +  
    scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2")) +
  labs(x = "Female Wing Chord", y = "Male Wing Chord") +
  theme_minimal()

```


## Breeding Success Analysis -- categorical wing difference (only male larger categories)

```{r}
head(breed)
breedCat <-breed[breed$Wing_Difference_Category %in% c("Much", "Slightly"), ]

#first egg, removing blank values
breedCat.fe <- breedCat %>% filter(!is.na(First.Egg))
sum(breedCat.fe$Species == "MOCH") 
sum(breedCat.fe$Species == "BCCH")

#clutch size, removing blank values
breedCat.cs <- breedCat %>% filter(!is.na(Egg_Number))
sum(breedCat.cs$Species == "MOCH") 
sum(breedCat.cs$Species == "BCCH")

#brood size, removing blank values
breedCat.bs <- breedCat %>% filter(!is.na(Nestling_Number))
sum(breedCat.bs$Species == "MOCH") 
sum(breedCat.bs$Species == "BCCH")

#nestling size, removing blank values
breedCat.mm <- breedCat %>% filter(!is.na(Avg_Nestling_Weight))
sum(breedCat.mm$Species == "MOCH") 
sum(breedCat.mm$Species == "BCCH")

#Female SMI, removing blank values
breedCat.fs <- breedCat %>% filter(!is.na(Female_SMI))
sum(breedCat.fs$Species == "MOCH") 
sum(breedCat.fs$Species == "BCCH")

#Male SMI, removing blank values
breedCat.ms <- breedCat %>% filter(!is.na(Male_SMI))
sum(breedCat.ms$Species == "MOCH") 
sum(breedCat.ms$Species == "BCCH")

#Provisioning
breedCat.p <- breedCat %>% filter(!is.na(Provisioning))
sum(breedCat.p$Species == "MOCH") 
sum(breedCat.p$Species == "BCCH")

```


## Modeling date of first egg

```{r}
# Centering within year
breedCat.fe$First.Egg_YearCentered <- ave(breedCat.fe$First.Egg, breedCat.fe$Year, FUN = function(x) scale(x, center = TRUE, scale = FALSE))

# Centering within elevation (after year)
breedCat.fe$First.Egg_Scaled <- ave(breedCat.fe$First.Egg_YearCentered, breedCat.fe$scale_ele, FUN = function(x) scale(x, center = TRUE, scale = FALSE))


fec1 <- lmer(First.Egg_Scaled ~ Wing_Difference_Category*Species + scale_ele + (1|Year), data = breedCat.fe)
fec2 <- lmer(First.Egg_Scaled ~ Species + scale_ele + (1|Year), data = breedCat.fe)

anova(fec1,fec2) #wing difference doesnt improve model fit

#Check residuals
fec1r = simulateResiduals(fec1)
plot(fec1r)
#strong deviation

summary(fec1)
catPFEB <-0.5776
catPFEM <-0.1059

```


## Modeling clutch size

```{r}
cs1 <- glmmTMB(Egg_Number ~ Wing_Difference_Category*Species, data =breedCat.cs,family = "genpois")
cs2 <- glmmTMB(Egg_Number ~ Species, data=breedCat.cs, family="genpois") #got error when including (1|Year) and variance of the random effect was very low

anova(cs1,cs2) # wing difference does not improve model fit

#Check residuals
cs1r = simulateResiduals(cs1)
plot(cs1r)
#look okay

summary(cs1)

catPBSB <- 0.206
catPBSM <-0.568

```

## Modeling brood size

```{r}
bs1 <- glmmTMB(Nestling_Number ~ Wing_Difference_Category*Species, data = breedCat.bs, family="genpois")
bs2 <- glmmTMB(Nestling_Number ~ Species, data = breedCat.bs, family = "genpois") #got error when including (1|Year) and variance of the random effect was very low

anova(bs1,bs2) #wing difference doesnt improve model fit

#check residuals
bs1r = simulateResiduals(bs1)
plot(bs1r)
#look ok

#Results
summary(bs1)

catPNNM <- 0.486
catPNNB <- 0.915

```

## Modeling average nestling mass

```{r}
mm3 <- lmer(Avg_Nestling_Weight ~ Wing_Difference_Category*Species + (1|Year), data = breedCat.mm)
mm4 <- lmer(Avg_Nestling_Weight ~ Species + (1|Year), data = breedCat.mm)

anova(mm3, mm4) #wing difference significantly improves model fit

#check residuals
mm3r = simulateResiduals(mm3)
plot(mm3r)
#looks good

#results
summary(mm3)

catPMMM <- 0.0259
catPMMB <- 0.715

```


## Modeling female SMI

```{r}
fs1 <- lmer(Female_SMI ~ Wing_Difference_Category*Species + (1|Year), data =  breedCat.fs)
fs2 <- lmer(Female_SMI ~ Species + (1|Year), data = breedCat.fs)

anova(fs1,fs2) #wing difference doesnt improve model fit

#check residuals
fs1r = simulateResiduals(fs1)
plot(fs1r)
#look okay

#results
summary(fs1) 

catPFSM <- 0.6
catPFSB <- 0.708

```


## Modeling male SMI
```{r}
ms1 <- lmer(Male_SMI ~ Wing_Difference_Category*Species + (1|Year), data =  breedCat.ms)
ms2 <- lmer(Male_SMI ~ Species + (1|Year), data = breedCat.ms)

anova(ms1,ms2) #wing difference doesnt improve model fit

#check residuals
ms1r = simulateResiduals(ms1)
plot(ms1r)
#look good

#results
summary(ms1) #wing_difference BCCH almost significant

catPMSM <- 0.3122
catPMSB <- 0.0765

```

## Modeling provisioning rate

```{r}
p1 <- lmer(Provisioning ~ Wing_Difference_Category*Species + (1|Year), data =  breedCat.p)
p2 <- lmer(Provisioning ~ Species + (1|Year), data = breedCat.p)

anova(p1,p2) #wing difference doesnt improve model fit

#check residuals
p1r = simulateResiduals(p1)
plot(p1r)
#quantile deviations

#results
summary(p1)

catPPM <- 0.469
catPPB <- 0.536

```
Adjust the p-values to account for multiple comparisons

```{r}
p_values <- c(catPFEM, catPFEB, catPBSM, catPBSB, catPNNM, catPNNB, catPMMM, catPMMB, catPFSM, catPFSB, catPMSM, catPMSB, catPPM, catPPB)

p_adjusted <- p.adjust(p_values, method = "BH")

p_adjusted

```


## Plotting wing difference category comparisons

```{r}
#moch first egg date ##
mochBreed.fe.plot <- subset(breed.fe, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxfe <- ggplot(data=mochBreed.fe.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=First.Egg)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Date of first egg")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#moch clutch size ##
mochBreed.cs.plot <- subset(breed.cs, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxcs <- ggplot(data=mochBreed.cs.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Egg_Number)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Clutch Size")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#moch brood size ##
mochBreed.bs.plot <- subset(breed.bs, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxbs <- ggplot(data=mochBreed.bs.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Nestling_Number)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Brood size")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))


#moch mean nestling mass ##
mochBreed.mm.plot <- subset(breed.mm, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxmm <-ggplot(data=mochBreed.mm.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Avg_Nestling_Weight)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Mean nestling mass")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#moch female SMI ##
mochBreed.fs.plot <- subset(breed.fs, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxfs <- ggplot(data=mochBreed.fs.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Female_SMI)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Female body condition")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))


#moch male SMI ##
mochBreed.ms.plot <- subset(breed.ms, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxms <- ggplot(data=mochBreed.ms.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Male_SMI)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Male body condition")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#moch provisioning ##
mochBreed.p.plot <- subset(breed.p, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxp <- ggplot(data=mochBreed.p.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Provisioning)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Provisioning")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

pMOCH <- (boxfe+boxcs+boxbs+boxmm)/(boxfs+boxms+boxp)

#bcch first egg date ##
bcchBreed.fe.plot <- subset(breed.fe, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxfebc <- ggplot(data=bcchBreed.fe.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=First.Egg)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Date of first egg")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#bcch clutch size ##
bcchBreed.cs.plot <- subset(breed.cs, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxcsbc <- ggplot(data=bcchBreed.cs.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Egg_Number)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Clutch Size")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#bcch brood size ##
bcchBreed.bs.plot <- subset(breed.bs, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxbsbc <- ggplot(data=bcchBreed.bs.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Nestling_Number)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Brood size")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))


#bcch mean nestling mass ##
bcchBreed.mm.plot <- subset(breed.mm, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxmmbc <-ggplot(data=bcchBreed.mm.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Avg_Nestling_Weight)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Mean nestling mass")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#bcch female SMI ##
bcchBreed.fs.plot <- subset(breed.fs, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxfsbc <- ggplot(data=bcchBreed.fs.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Female_SMI)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Female body condition")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))


#bcch male SMI ##
bcchBreed.ms.plot <- subset(breed.ms, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxmsbc <- ggplot(data=bcchBreed.ms.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Male_SMI)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Male body condition")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#bcch provisioning ##
bcchBreed.p.plot <- subset(breed.p, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxpbc <- ggplot(data=bcchBreed.p.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Provisioning)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Provisioning")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

pBCCH <- (boxfebc+boxcsbc+boxbsbc+boxmmbc)/(boxfsbc+boxmsbc+boxpbc)

ggsave("boxplotMOCH.png", plot = pMOCH, width = 7, height = 5, units = "in", dpi = 300)
ggsave("boxplotBCCH.png", plot = pBCCH, width = 7, height = 5, units = "in", dpi = 300)


```


```{r}


library(ggeffects)

preds <- ggpredict(cs1, terms = c("Wing_Difference", "Species"))

ggplot(preds, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
  labs(x = "Wing Length Difference", y = "Predicted Clutch Size", color = "Species", fill = "Species") +
  theme_minimal()

library(ggeffects)
preds <- ggpredict(bs1, terms = c("Wing_Difference", "Species"))

ggplot(preds, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
  labs(x = "Wing Length Difference", y = "Predicted Nestling Number", color = "Species", fill = "Species") +
  theme_minimal()


```