---
title: "Padula_2025_Wing_Diff_Analysis"
author: "Sara Padula"
date: "2025-03-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

```{r}
library(dplyr)
library(ggplot2)
library(cowplot)
library(lme4)
library(lmerTest)
library(DHARMa)
library(glmmTMB)
library(car)
library(emmeans)
library(ggpubr)
library(ggdist)
library(patchwork)
```

## Load data


```{r}
#Loading data
breed <- read.csv("paired_SMI.csv")

## Scaling elevation variable for each species
breedMOCH <- breed %>%
  filter(Species == "MOCH") %>%
  mutate(scale_ele = scale(Elevation))

breedBCCH <- breed %>%
  filter(Species == "BCCH") %>%
  filter(Elevation < 2000) %>%
  mutate(scale_ele = scale(Elevation))

breed <- rbind(breedMOCH, breedBCCH)

#What is the sample size and mean wing difference in...

#MOCH?
nrow(breedMOCH)

mean(breedMOCH$Wing_Difference)


#BCCH?
nrow(breedBCCH)

mean(breedBCCH$Wing_Difference)

```

## Breeding Success Analysis -- Numerical wing difference variable
```{r}
#first egg, removing blank values
breed.fe <- breed %>% filter(!is.na(First.Egg))
sum(breed.fe$Species == "MOCH") 
sum(breed.fe$Species == "BCCH")

#clutch size, removing blank values
breed.cs <- breed %>% filter(!is.na(Egg_Number))
sum(breed.cs$Species == "MOCH") 
sum(breed.cs$Species == "BCCH")

#brood size, removing blank values
breed.bs <- breed %>% filter(!is.na(Nestling_Number))
sum(breed.bs$Species == "MOCH") 
sum(breed.bs$Species == "BCCH")

#nestling size, removing blank values
breed.mm <- breed %>% filter(!is.na(Avg_Nestling_Weight))
sum(breed.mm$Species == "MOCH") 
sum(breed.mm$Species == "BCCH")

#Female SMI, removing blank values
breed.fs <- breed %>% filter(!is.na(Female_SMI))
sum(breed.fs$Species == "MOCH") 
sum(breed.fs$Species == "BCCH")

#Male SMI, removing blank values
breed.ms <- breed %>% filter(!is.na(Male_SMI))
sum(breed.ms$Species == "MOCH") 
sum(breed.ms$Species == "BCCH")

#Provisioning
breed.p <- breed %>% filter(!is.na(Provisioning))
sum(breed.p$Species == "MOCH") 
sum(breed.p$Species == "BCCH")


```

## Modeling date of first egg

```{r}
# Centering within year
breed.fe$First.Egg_YearCentered <- ave(breed.fe$First.Egg, breed.fe$Year, FUN = function(x) scale(x, center = TRUE, scale = FALSE))

# Centering within elevation (after year)
breed.fe$First.Egg_Scaled <- ave(breed.fe$First.Egg_YearCentered, breed.fe$scale_ele, FUN = function(x) scale(x, center = TRUE, scale = FALSE))


## Model first egg ######
fe1 <- lmer(First.Egg_Scaled ~ Wing_Difference*Species + Species*scale_ele + (1|Year), data = breed.fe)
fe2 <- lmer(First.Egg_Scaled ~ Wing_Difference*Species + scale_ele + (1|Year), data = breed.fe)

AIC(fe1,fe2) #fe1 lower AIC
anova(fe1,fe2) #not significant

#Check residuals
fe2r = simulateResiduals(fe2)
plot(fe2r)
#quantile deviation

Anova(fe1,test.statistic = "Chisq",type="III")
summary(fe1) # species significant

mochFEp <- 0.5403

bcchFEp <- 0.2873


```

## Modeling clutch size

```{r}
cs1 <- glmmTMB(Egg_Number ~ Wing_Difference*Species + Species*scale_ele, data =breed.cs,family = "genpois")
cs2 <- glmmTMB(Egg_Number ~ Wing_Difference*Species + scale_ele, data=breed.cs, family="genpois") #got error when including (1|Year) and variance of the random effect was very low

AIC(cs1, cs2) #cs2 lower AIC
anova(cs1,cs2) # not significant

#Check residuals
cs2r = simulateResiduals(cs2)
plot(cs2r)
#look okay

Anova(cs2,test.statistic = "Chisq",type="III")
summary(cs2) #species significant

mochCSp <- 0.118

bcchCSp <- 0.738


```

## Modeling brood size

```{r}
bs1 <- glmmTMB(Nestling_Number ~ Wing_Difference*Species + Species*scale_ele, data = breed.bs, family="genpois")
bs2 <- glmmTMB(Nestling_Number ~ Wing_Difference*Species + scale_ele, data = breed.bs, family = "genpois") #got error when including (1|Year) and variance of the random effect was very low

AIC(bs1,bs2) #bs2 lower AIC
anova(bs1,bs2) #not significant

#check residuals
bs2r = simulateResiduals(bs2)
plot(bs2r)
#look good

#Results
Anova(bs2,test.statistic = "Chisq",type="III")
summary(bs2) #species and wing_difference*species significant

mochBSp <- 0.00113

bcchBSp <- 0.30050


```

## Modeling average nestling mass

```{r}
##Model nestling mass #####
mm1 <- lmer(Avg_Nestling_Weight ~ Wing_Difference*Species + Species*scale_ele + Nestling_Number + (1|Year), data = breed.mm)
mm2 <- lmer(Avg_Nestling_Weight ~ Wing_Difference*Species + scale_ele + Nestling_Number + (1|Year), data = breed.mm)
mm3 <- lmer(Avg_Nestling_Weight ~ Wing_Difference*Species + Species*scale_ele + (1|Year), data = breed.mm)
mm4 <- lmer(Avg_Nestling_Weight ~ Wing_Difference*Species + scale_ele + (1|Year), data = breed.mm)

AIC(mm1, mm2, mm3, mm4) #mm3 lowest AIC

#check residuals
mm3r = simulateResiduals(mm3)
plot(mm3r)
#some deviation

#results
Anova(mm3,test.statistic = "Chisq",type="III")
summary(mm3) #wing_difference*species significant

mochMMp <- 0.0119

bcchMMp <- 0.8479
```

## Modeling Female SMI

```{r}
fs1 <- lmer(Female_SMI ~ Wing_Difference*Species + Species*scale_ele + (1|Year), data =  breed.fs)
fs2 <- lmer(Female_SMI ~ Wing_Difference*Species + scale_ele + (1|Year), data = breed.fs)

AIC(fs1, fs2) #fs1 lower AIC
anova(fs1,fs2) #significant

#check residuals
fs1r = simulateResiduals(fs1)
plot(fs1r)
#look okay

#results
Anova(fs1,test.statistic = "Chisq",type="III")
summary(fs1) #species*elevation significant, elevation almost significant

mochFSp <- 0.86336

bcchFSp <- 0.42181

```

## Modeling male SMI

```{r}
ms1 <- lmer(Male_SMI ~ Wing_Difference*Species + Species*scale_ele + (1|Year), data =  breed.ms)
ms2 <- lmer(Male_SMI ~ Wing_Difference*Species + scale_ele + (1|Year), data = breed.ms)

AIC(ms1, ms2) #ms2 lower AIC
anova(ms1,ms2) #not significant

#check residuals
ms2r = simulateResiduals(ms2)
plot(ms2r)
#look good

#results
Anova(ms2,test.statistic = "Chisq",type="III")
summary(ms2) #wing_difference significant

mochMSp <- 0.15327

bcchMSp <- 0.00642
```

## Modeling provisioning rate

```{r}
p1 <- lmer(Provisioning ~ Wing_Difference*Species + Species*scale_ele + (1|Year), data =  breed.p)
p2 <- lmer(Provisioning ~ Wing_Difference*Species + scale_ele + (1|Year), data = breed.p)

AIC(p1, p2) #p1 lower AIC
anova(p1,p2) #not significant

#check residuals
p1r = simulateResiduals(p1)
plot(p1r)
#quantile deviations

#results
Anova(p1,test.statistic = "Chisq",type="III")
summary(p1) #nothing significant

mochPp <- 0.695

bcchPp <- 0.966
```

Adjust the p-values to account for multiple comparisons

```{r}
p_values <- c(mochFEp, bcchFEp, mochCSp, bcchCSp, mochBSp, bcchBSp, mochMMp, bcchMMp, mochFSp, bcchFSp, mochMSp, bcchMSp, mochPp, bcchPp)

p_adjusted <- p.adjust(p_values, method = "BH")

p_adjusted

```


Visualizing wing difference (numeric) predicting reproductive success variables.

```{r}
a <- ggplot(breed.fe, aes(x = Wing_Difference, y = First.Egg_Scaled, color = Species)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2")) +  
  labs(x = "",
       y = "Scaled First Egg Date") +
  theme_minimal()+ 
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 7))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

b <-ggplot(breed.cs, aes(x = Wing_Difference, y = Egg_Number, color = Species)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2")) +  
  labs(x = "",
       y = "Clutch Size") +
  theme_minimal()+
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 7))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

c <- ggplot(breed.bs, aes(x = Wing_Difference, y = Nestling_Number, color = Species)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2")) +  
    labs(x = "",
       y = "Brood Size")+
    theme_minimal()+
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 7))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

d <- ggplot(breed.mm, aes(x = Wing_Difference, y = Avg_Nestling_Weight, color = Species)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2")) +  
  labs(x = "",
       y = "Average Nestling Mass (g)") +
  theme_minimal()+
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 7))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

e <- ggplot(breed.fs, aes(x = Wing_Difference, y = Female_SMI, color = Species)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2")) +  
  labs(x = "",
       y = "Female Body Condition") +
  theme_minimal()+
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 7))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

f <- ggplot(breed.ms, aes(x = Wing_Difference, y = Male_SMI, color = Species)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2")) +  
  labs(x = "Wing Length Difference (mm)",
       y = "Male Body Condition") +
  theme_minimal()+
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 7))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

g <- ggplot(breed.p, aes(x = Wing_Difference, y = Provisioning, color = Species)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2")) +  
  labs(x = "Wing Length Difference (mm)",
       y = "Provisioning rate") +
  theme_minimal()+
    theme(legend.position = "none")+
    theme(axis.title.y = element_text(size = 7))+
    theme(axis.text.y = element_text(size = 5))+
    theme(axis.text.x = element_text(size = 5))

allLine <- (a + b + c)/(d + e + f + g)

ggsave("wingdiffRepVarPlot.png", plot = allLine, width = 6, height = 4, dpi = 300)

```


## Does female wing length predict male wing length within pairs?

```{r}
#subsetting data
mochBreed <- subset(breed, Species == "MOCH")
bcchBreed <- subset(breed, Species == "BCCH")

#create new variable for female wing chord

mochBreed$Female.Wing.Chord2 <- mochBreed$Female.Wing.Chord^2

bcchBreed$Female.Wing.Chord2 <- bcchBreed$Female.Wing.Chord^2

#MOCH

#which quadratic model is the best?
head(mochBreed)
quadM1 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + Female.Wing.Chord2 + FemaleBander + MaleBander + Year, data = mochBreed)
quadM2 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + Female.Wing.Chord2 + FemaleBander + MaleBander, data = mochBreed)
quadM3 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + Female.Wing.Chord2 + Year, data = mochBreed)
quadM4 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + Female.Wing.Chord2, data = mochBreed)

anova(quadM1, quadM2, quadM3, quadM4, test = "F")

#none of the additional factors improve model fit, the simplest model fits just as well

#which linear model is the best?
head(mochBreed)
lmM1 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + FemaleBander + MaleBander + Year, data = mochBreed)
lmM2 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + FemaleBander + MaleBander, data = mochBreed)
lmM3 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + Year, data = mochBreed)
lmM4 <- lm(Male.Wing.Chord ~ Female.Wing.Chord, data = mochBreed)

anova(lmM1, lmM2, lmM3, lmM4, test = "F")

#adding predictors does not improve model fit, the simplest model fits just as well

quadraticMoch <- lm(Male.Wing.Chord ~ Female.Wing.Chord + Female.Wing.Chord2, data = mochBreed)
lmMoch <- lm(Male.Wing.Chord ~ Female.Wing.Chord, data = mochBreed)
nullMoch <- lm(Male.Wing.Chord ~ 1, data = mochBreed)

summary(quadraticMoch)
summary(lmMoch)

AIC(quadraticMoch,lmMoch) 

#quadratic has a higher AIC score but not high enough to justify doing the more complex model

anova(nullMoch,lmMoch)

#the model does not explain a meaningful amount of variation 

#BCCH

#Does female wing predict male wing in BCCH

#which quadratic model is the best?
head(bcchBreed)
quadB1 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + Female.Wing.Chord2 + FemaleBander + MaleBander + Year, data = bcchBreed)
quadB2 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + Female.Wing.Chord2 + FemaleBander + MaleBander, data = bcchBreed)
quadB3 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + Female.Wing.Chord2 + Year, data = bcchBreed)
quadB4 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + Female.Wing.Chord2, data = bcchBreed)

anova(quadB1, quadB2, quadB3, quadB4, test = "F")

#male bander and female bander are essential, quadB2 is the best model fit

#which linear model is the best?
head(bcchBreed)
lmB1 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + FemaleBander + MaleBander + Year, data = bcchBreed)
lmB2 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + FemaleBander + MaleBander, data = bcchBreed)
lmB3 <- lm(Male.Wing.Chord ~ Female.Wing.Chord + Year, data = bcchBreed)
lmB4 <- lm(Male.Wing.Chord ~ Female.Wing.Chord, data = bcchBreed)

anova(lmB1, lmB2, lmB3, lmB4, test = "F")

#male bander and female bander are essential, lmB2 is the best model fit

quadraticBcch <- lm(Male.Wing.Chord ~ Female.Wing.Chord + Female.Wing.Chord2, data = bcchBreed)
lmBcch <- lm(Male.Wing.Chord ~ Female.Wing.Chord, data = bcchBreed)
nullBcch <- lm(Male.Wing.Chord ~ FemaleBander + MaleBander, data = bcchBreed)

summary(quadraticBcch)
summary(lmBcch)

AIC(quadraticBcch,lmBcch) 

#quadratic has a higher AIC score but not high enough to justify doing the more complex model

anova(nullBcch,lmBcch)

#the model does not explain a meaningful amount of variation 

```

# Plot the linear models
```{r}

ggplot(breed, aes(x = Female.Wing.Chord, y = Male.Wing.Chord, color = Species)) +
  geom_point(alpha = 0.7) +  
  geom_smooth(method = "lm", se = FALSE) +  
    scale_color_manual(values = c("MOCH" = "#cc79a7", "BCCH" = "#0072b2")) +
  labs(x = "Female Wing Chord", y = "Male Wing Chord") +
  theme_minimal()

```


## Breeding Success Analysis -- categorical wing difference (only male larger categories)

```{r}
head(breed)
breedCat <-breed[breed$Wing_Difference_Category %in% c("Much", "Slightly"), ]

#first egg, removing blank values
breedCat.fe <- breedCat %>% filter(!is.na(First.Egg))
sum(breedCat.fe$Species == "MOCH") 
sum(breedCat.fe$Species == "BCCH")

#clutch size, removing blank values
breedCat.cs <- breedCat %>% filter(!is.na(Egg_Number))
sum(breedCat.cs$Species == "MOCH") 
sum(breedCat.cs$Species == "BCCH")

#brood size, removing blank values
breedCat.bs <- breedCat %>% filter(!is.na(Nestling_Number))
sum(breedCat.bs$Species == "MOCH") 
sum(breedCat.bs$Species == "BCCH")

#nestling size, removing blank values
breedCat.mm <- breedCat %>% filter(!is.na(Avg_Nestling_Weight))
sum(breedCat.mm$Species == "MOCH") 
sum(breedCat.mm$Species == "BCCH")

#Female SMI, removing blank values
breedCat.fs <- breedCat %>% filter(!is.na(Female_SMI))
sum(breedCat.fs$Species == "MOCH") 
sum(breedCat.fs$Species == "BCCH")

#Male SMI, removing blank values
breedCat.ms <- breedCat %>% filter(!is.na(Male_SMI))
sum(breedCat.ms$Species == "MOCH") 
sum(breedCat.ms$Species == "BCCH")

#Provisioning
breedCat.p <- breedCat %>% filter(!is.na(Provisioning))
sum(breedCat.p$Species == "MOCH") 
sum(breedCat.p$Species == "BCCH")

```


## Modeling date of first egg

```{r}
# Centering within year
breedCat.fe$First.Egg_YearCentered <- ave(breedCat.fe$First.Egg, breedCat.fe$Year, FUN = function(x) scale(x, center = TRUE, scale = FALSE))

# Centering within elevation (after year)
breedCat.fe$First.Egg_Scaled <- ave(breedCat.fe$First.Egg_YearCentered, breedCat.fe$scale_ele, FUN = function(x) scale(x, center = TRUE, scale = FALSE))


fec1 <- lmer(First.Egg_Scaled ~ Wing_Difference_Category*Species + Species*scale_ele + (1|Year), data = breedCat.fe)
fec2 <- lmer(First.Egg_Scaled ~ Wing_Difference_Category*Species + scale_ele + (1|Year), data = breedCat.fe)

AIC(fec1,fec2) #fe1 lower AIC
anova(fec1,fec2) #significant

#Check residuals
fec1r = simulateResiduals(fec1)
plot(fec1r)
#looks good

Anova(fec1,test.statistic = "Chisq",type="III")
summary(fec1) #species, elevation, and species*elevation significant


```


## Modeling clutch size

```{r}
cs1 <- glmmTMB(Egg_Number ~ Wing_Difference_Category*Species + Species*scale_ele, data =breedCat.cs,family = "genpois")
cs2 <- glmmTMB(Egg_Number ~ Wing_Difference_Category*Species + scale_ele, data=breedCat.cs, family="genpois") #got error when including (1|Year) and variance of the random effect was very low

AIC(cs1, cs2) #cs2 lower AIC
anova(cs1,cs2) # not significant

#Check residuals
cs2r = simulateResiduals(cs2)
plot(cs2r)
#look okay

Anova(cs2,test.statistic = "Chisq",type="III")
summary(cs2) #species significant

```

## Modeling brood size

```{r}
bs1 <- glmmTMB(Nestling_Number ~ Wing_Difference_Category*Species + Species*scale_ele, data = breedCat.bs, family="genpois")
bs2 <- glmmTMB(Nestling_Number ~ Wing_Difference_Category*Species + scale_ele, data = breedCat.bs, family = "genpois") #got error when including (1|Year) and variance of the random effect was very low

AIC(bs1,bs2) #bs2 lower AIC
anova(bs1,bs2) #not significant

#check residuals
bs2r = simulateResiduals(bs2)
plot(bs2r)
#look ok

#Results
Anova(bs2,test.statistic = "Chisq",type="III")
summary(bs1) #nothing significant

```

## Modeling average nestling mass

```{r}
mm3 <- lmer(Avg_Nestling_Weight ~ Wing_Difference_Category*Species + Species*scale_ele + (1|Year), data = breedCat.mm)
mm4 <- lmer(Avg_Nestling_Weight ~ Wing_Difference_Category*Species + scale_ele + (1|Year), data = breedCat.mm)

AIC(mm3, mm4) #mm3 lowest AIC
anova( mm3, mm4) #significant

#check residuals
mm3r = simulateResiduals(mm3)
plot(mm3r)
#some deviation

#results
Anova(mm3,test.statistic = "Chisq",type="III")
summary(mm3) #wing_difference*species significant

```


## Modeling female SMI

```{r}
fs1 <- lmer(Female_SMI ~ Wing_Difference_Category*Species + Species*scale_ele + (1|Year), data =  breedCat.fs)
fs2 <- lmer(Female_SMI ~ Wing_Difference_Category*Species + scale_ele + (1|Year), data = breedCat.fs)

AIC(fs1, fs2) #fs1 lower AIC
anova(fs1,fs2) #significant

#check residuals
fs1r = simulateResiduals(fs1)
plot(fs1r)
#look okay

#results
Anova(fs1,test.statistic = "Chisq",type="III")
summary(fs1) #species*elevation significant

```


## Modeling male SMI
```{r}
ms1 <- lmer(Male_SMI ~ Wing_Difference_Category*Species + Species*scale_ele + (1|Year), data =  breedCat.ms)
ms2 <- lmer(Male_SMI ~ Wing_Difference_Category*Species + scale_ele + (1|Year), data = breedCat.ms)

AIC(ms1, ms2) #ms2 lower AIC
anova(ms1,ms2) #not significant

#check residuals
ms2r = simulateResiduals(ms2)
plot(ms2r)
#look good

#results
Anova(ms2,test.statistic = "Chisq",type="III")
summary(ms2) #wing_difference almost significant

```

## Modeling provisioning rate

```{r}
p1 <- lmer(Provisioning ~ Wing_Difference_Category*Species + Species*scale_ele + (1|Year), data =  breedCat.p)
p2 <- lmer(Provisioning ~ Wing_Difference_Category*Species + scale_ele + (1|Year), data = breedCat.p)

AIC(p1, p2) #p1 lower AIC
anova(p1,p2) #not significant

#check residuals
p1r = simulateResiduals(p1)
plot(p1r)
#quantile deviations

#results
Anova(p1,test.statistic = "Chisq",type="III")
summary(p1) #nothing significant

```


## Plotting wing difference category comparisons

```{r}
#moch first egg date ##
mochBreed.fe.plot <- subset(breed.fe, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxfe <- ggplot(data=mochBreed.fe.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=First.Egg)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Day of first egg")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#moch clutch size ##
mochBreed.cs.plot <- subset(breed.cs, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxcs <- ggplot(data=mochBreed.cs.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Egg_Number)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Clutch Size")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#moch brood size ##
mochBreed.bs.plot <- subset(breed.bs, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxbs <- ggplot(data=mochBreed.bs.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Nestling_Number)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Brood size")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))


#moch mean nestling mass ##
mochBreed.mm.plot <- subset(breed.mm, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxmm <-ggplot(data=mochBreed.mm.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Avg_Nestling_Weight)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Mean nestling mass")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#moch female SMI ##
mochBreed.fs.plot <- subset(breed.fs, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxfs <- ggplot(data=mochBreed.fs.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Female_SMI)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Female body condition")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))


#moch male SMI ##
mochBreed.ms.plot <- subset(breed.ms, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxms <- ggplot(data=mochBreed.ms.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Male_SMI)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Male body condition")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#moch provisioning ##
mochBreed.p.plot <- subset(breed.p, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("MOCH"))

boxp <- ggplot(data=mochBreed.p.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Provisioning)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Provisioning")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

(boxfe+boxcs+boxbs+boxmm)/(boxfs+boxms+boxp)

#bcch first egg date ##
bcchBreed.fe.plot <- subset(breed.fe, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxfebc <- ggplot(data=bcchBreed.fe.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=First.Egg)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Day of first egg")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#bcch clutch size ##
bcchBreed.cs.plot <- subset(breed.cs, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxcsbc <- ggplot(data=bcchBreed.cs.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Egg_Number)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Clutch Size")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#bcch brood size ##
bcchBreed.bs.plot <- subset(breed.bs, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxbsbc <- ggplot(data=bcchBreed.bs.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Nestling_Number)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Brood size")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))


#bcch mean nestling mass ##
bcchBreed.mm.plot <- subset(breed.mm, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxmmbc <-ggplot(data=bcchBreed.mm.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Avg_Nestling_Weight)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Mean nestling mass")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#bcch female SMI ##
bcchBreed.fs.plot <- subset(breed.fs, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxfsbc <- ggplot(data=bcchBreed.fs.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Female_SMI)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Female body condition")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))


#bcch male SMI ##
bcchBreed.ms.plot <- subset(breed.ms, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxmsbc <- ggplot(data=bcchBreed.ms.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Male_SMI)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Male body condition")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

#bcch provisioning ##
bcchBreed.p.plot <- subset(breed.p, Wing_Difference_Category %in% c("Slightly", "Much")) %>%
  subset(Species %in% c("BCCH"))

boxpbc <- ggplot(data=bcchBreed.p.plot,aes(x=factor(Wing_Difference_Category, levels = c("Slightly","Much")),y=Provisioning)) +
  geom_boxplot(aes(fill=Wing_Difference_Category),outlier.alpha = 0) +
  geom_point(position=position_jitter(width=0.1,height=0),alpha=0.6) +
  theme_cowplot() + theme(legend.position = "") + scale_fill_manual(values=c("gray","orange")) +
  xlab("") + ylab("Provisioning")+
  theme(axis.title = element_text(size = 10))+
  theme(axis.title.x = element_text(margin = margin(t = 8)))+
  theme(axis.title.y = element_text(margin = margin(r = 8)))+
  theme(axis.text = element_text(size=6, color = "black"))

(boxfebc+boxcsbc+boxbsbc+boxmmbc)/(boxfsbc+boxmsbc+boxpbc)

```